Precise predictions for same-sign W-boson scattering at the LHC

### 2018-02-21 Draft/acknowledgements.tex
Small corrections and Conclusion drafted

BB, AD, and MP acknowledge financial support by the
German Federal Ministry for Education and Research (BMBF) under
contract no.~05H15WWCA1 and the German Science Foundation (DFG) under
reference number DE 623/6-1.
%
AK acknowledges financial support by the Swiss National Science Foundation (SNF) under contract 200020-175595.
MR acknowledges funding from the European Union's Horizon 2020 research and innovation programme as part of the Marie Sklodowska-Curie Innovative Training Network MCnetITN3 (grant agreement no. 722104).
MZ is supported by the Netherlands National Organisation for Scientific Research (NWO),

>>>>

BB, AD, and MP acknowledge financial support by the
German Federal Ministry for Education and Research (BMBF) under
contract no.~05H15WWCA1 and the German Science Foundation (DFG) under
reference number DE 623/6-1.
AK acknowledges financial support by the Swiss National Science Foundation (SNF) under contract 200020-175595.
MR acknowledges funding from the European Union's Horizon 2020 research and innovation programme as part of the Marie Sklodowska-Curie Innovative Training Network MCnetITN3 (grant agreement no. 722104).
CS acknowledges support by the state of Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG) through grant no INST 39/963-1 FUGG and grant DI 784/3.
MZ is supported by the Netherlands National Organisation for Scientific Research (NWO).

### 2018-02-21 Draft/conclusion.tex
Small corrections and Conclusion drafted

In the present article, a detailed study of the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC has been presented.
The main focus is the electroweak (EW) production of such a final state where vector-boson scattering (VBS) occurs.
So far, all NLO computations have been performed in the so-called VBS approximation.
Only recently a full computation became available \cite{Biedermann:2017bss}.
Various theoretical predictions have been compared to the full computation.
This has not only been performed in a typical VBS fiducial region but also in more inclusive volumes.
We quantify precisely the differences that arise for several physical observables and in particular for the di-jet invariant mass and the rapidity-separation of the leading two jets.
This is the first time that such an in-depth study is performed and should be representative of the quality of the approximations for other VBS signatures.
In addition to fixed order predictions we have also investigated the impact of parton-shower.
To that end, several fixed-order computations matched to parton shower have been used.
It turns out that very large differences can appear for certain observables.
These differences appear in the central region where for VBS, colour-recombination plays a significant role.
These findings are new and should trigger further investigations in the theoretical community as well as by experimental collaborations.
Indeed, measurements of such observables should allow for a better understanding of colour-recombination in parton shower.
\MP{It might be too strong a statement. At least Simon will rephrase it.}
The results presented here are exclusively theoretical.
Nonetheless, they should raise significant interests in the experimental collaborations.
Therefore, to supplement this summary, we provide several recommendations for the use of theoretical predictions for the quest of measuring VBS processes precisely.

\MP{This might deserve a section on its own.} \\
Recommendations to experimental collaborations: \\
- Stress that this is only for W+W+ but that qualitative results should apply to other VBS signatures with massive gauge bosons.
- Combinations with EW NLO corrections. \\
- Missing higher EW order: $\pm \delta^2_{\rm NLO EW}$ \\
- Systematics when using NLO QCD approximation \\
- Systematics of different parton shower \\
- Combined measurement including  EW, QCD, and interference \\
- Move to NLO predictions / generators \\
- Comment on the irreducible QCD background \\
- Uncertainties related to PDF.
Some of us have already presented preliminary results on this subjetc \cite{}. % Radcor proceedings from Christopher.
A forthcoming article will addresses related questions.

>>>>

In the present article, a detailed study of the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC has been presented.
The main focus is the electroweak (EW) production of such a final state where vector-boson scattering (VBS) occurs.
So far, all NLO computations have been performed in the so-called VBS approximation.
Only recently a full computation became available \cite{Biedermann:2017bss}.
Various theoretical predictions have been compared to the full computation.
This has not only been performed in a typical VBS fiducial region but also in more inclusive volumes.
We quantify precisely the differences that arise for several physical observables and in particular for the di-jet invariant mass and the rapidity-separation of the leading two jets.
This is the first time that such an in-depth study is performed and should be representative of the quality of the approximations for other VBS signatures.
In addition to fixed order predictions we have also investigated the impact of parton-shower.
To that end, several fixed-order computations matched to parton shower have been used.
It turns out that very large differences can appear for certain observables.
These differences appear in the central region where for VBS, colour-recombination plays a significant role.
These findings are new and should trigger further investigations in the theoretical community as well as by experimental collaborations.
Indeed, measurements of such observables should allow for a better understanding of colour-recombination in parton shower.
\MP{It might be too strong a statement. At least Simon will rephrase it.}

The results presented here are exclusively theoretical.
Nonetheless, they should raise significant interests in the experimental collaborations.
Therefore, to supplement this summary, we provide several concluding remarks concerning the use of theoretical predictions.

As found in Ref.~\cite{Biedermann:2017bss}, the NLO EW corrections of order $\mathcal{O}{\left(\alpha^{7}\right)}$ are the dominant NLO contributions to the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$.
It is thus highly desirable that they are combined with NLO-QCD predictions matched with parton shower.
In Ref.~\cite{Biedermann:2016yds}, it has been demonstrated that these large EW corrections originate from the Sudakov logarithms which factorise.
Therefore we recommend to combined them in a multiplicative way.
Concerning the estimation of missing higher-order corrections of EW type they can be obtained, in a first approximation, from $\pm \delta^2_{\rm NLO EW}$.
The missing higher-order mixed QCD-EW corrections can be estimate by taking the difference between the multiplicative and additive prescription.
A detailed study of the combination of higher-order corrections has been carried out in the context of SM background for dark matter searches at the LHC \cite{Lindert:2017olm} and could be followed here as well.

We have shown how good the VBS approximation is performing in both inclusive and exclusive set-ups at both LO and NLO-QCD accuracy.
It turns out that for typical fiducial region used by experimental collaborations for their measurements, the agreement between the approximations and full calculation is certainly satisfactory given the current experimental precision.
Nonetheless, in the future when experimental measurements will improve, care has to be taken when using such approximations, in particular, if more inclusive phase-space are used.

Concerning the fixed-order predictions supplemented with parton shower, we have shown that for observables defined at LO, the differences are reasonable.
On the other hand, for observables only defined at NLO-QCD, such as the ones related to the third hardest jet, large differences can arise between different parton-shower algorithms.
Therefore, jet-veto procedures or selections based on extra radiation should be avoided as they carry rather large uncertainties.

In addition to the standard interpretation of EW signal versus QCD background, combined measurements should also be presented as they are well defined theoretically.
At LO, the interferences can be included in the background component.
Nonetheless, at NLO the concept of EW and QCD component become meaningless as at the order $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ both types of amplitudes mix.
Therefore, a combined measurement including the EW, QCD, and interference contributions is desirable.
Note that with such a measurement, an interpretation in terms of EW component is still possible.
This would amount to replace the subtraction of the QCD component based on Monte Carlo predictions by an actual measurement.
\MP{I am aware that this might be the most touchy point.
Therefore we should probably discuss it all together to ensure that we all agree on a definite statement.}

As demonstrated above, numerous programs exist to simulate the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC.
So far there is no public code that perform the full NLO computation matched with parton shower.
Nonetheless, full computations at NLO and publicly available tools at NLO-QCD accuracy in the VBS approximation matched to parton shower are available.
Therefore, we encourage experimental collaborations to move to NLO accuracy supplemented with parton shower as much as possible.
In particular, we are willing to support this effort. \MP{Are we?}

The present study has focused on the orders $\mathcal{O}{\left(\alpha^{6}\right)}$ at LO and $\mathcal{O}{\left(\alphas\alpha^{6}\right)}$ at NLO.
Nonetheless NLO computations and publicly available tools also exist for the QCD-induced process \cite{Rauch:2016pai,Melia:2010bm,Melia:2011gk,Campanario:2013gea,Baglio:2014uba,Biedermann:2017bss}.
They should therefore be also used as much as possible.

Uncertainties related to PDF has been hardly addressed in the present article.
It has only be discussed briefly in Sec.~\ref{sec:matching} where scale variation in combination with PDF variation has been shown.
Some of us have already presented preliminary results on this subject \cite{Schwan:2017yy}. % Radcor proceedings from Christopher.
Forthcoming publications will thus address related questions.

Finally, the study presented here concentrates on the ${\rm W}^+{\rm W}^+$ signature.
The features observed here (validity of the VBS approximation or comparison of theoretical predictions matched to parton shower for example) should be qualitatively similar for other VBS signatures with massive gauge bosons.
But the quantitative conclusions are strictly restricted to ${\rm W}^+{\rm W}^+$ signature.
For other signatures, similar quantitative study should be performed.
Beyond the theoretical findings of this work, we also hope that it can serve as a guideline for experimental collaborations in their quest for the measurements of VBS processes at the LHC.

### 2018-02-21 Draft/scanNLO.tex
Small corrections and Conclusion drafted

The same positive discrepancy for the $|s|^2 + |t|^2 + |u|^2$ approximation, can be seen in the low $\pt$ region of the leading jet in the upper plot of Fig.~\ref{fig:mjjdyjj_1d_2}.
In the large invariant mass--small rapidity separation region of Fig.~\ref{fig:ratio2d_NLO}, discrepancies at the level of $15\%$ are present.
This can be traced back to the large $\pt$ and central rapidity region of the leading jets kinematics, shown in Fig.~\ref{fig:mjjdyjj_1d_2}.
For such distributions, despite the $s$-channel inclusion, the discrepancy between the approximated and full result is about $5-10\%$.
In the VBS signal-region the VBS approximation shows a good agreement with the full calculation as documented in details below.

>>>>

The same positive discrepancy for the $|s|^2 + |t|^2 + |u|^2$ approximation, can be seen in the low transverse-momentum region of the leading jet in the upper plot of Fig.~\ref{fig:mjjdyjj_1d_2}.
In the large invariant mass--small rapidity separation region of Fig.~\ref{fig:ratio2d_NLO}, discrepancies at the level of $15\%$ are present.
This can be traced back to the large $\pt$ and central rapidity region of the leading jets kinematics, shown in Fig.~\ref{fig:mjjdyjj_1d_2}.
For such distributions, despite the $s$-channel inclusion, the discrepancy between the approximated and full result is about $5-10\%$.
In the VBS signal-region the VBS approximation shows a good agreement with the full calculation as documented in details below.

### 2018-02-21 Draft/shower.tex
Small corrections and Conclusion drafted

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. 
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated, e.g. when fully-differential computations at NNLO will become available (for VBF Higgs production, see Refs.~\cite{Cacciari:2015jma, Cruz-Martinez:2018rod}. It is a remarkable fact that, even for those observables that display small discrepancies,
the theoretical uncertainty obtained via scale variations systematically underestimates the spread of predictions. Again, this stresses the need
 to employ at least two different tools in order to have a conservative estimate of theoretical uncertainties. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as little as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties.\AK{Are there any current analyses with a central jet veto in VBS? The experimentalists might take this statement a little too literal.} On the one hand, in order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable; on the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
 VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data. \AK{Include some references here?}

>>>>

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. 
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated, e.g. when fully-differential computations at NNLO will become available (for VBF Higgs production, see Refs.~\cite{Cacciari:2015jma, Cruz-Martinez:2018rod}). It is a remarkable fact that, even for those observables that display small discrepancies,
the theoretical uncertainty obtained via scale variations systematically underestimates the spread of predictions. Again, this stresses the need
 to employ at least two different tools in order to have a conservative estimate of theoretical uncertainties. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as little as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties.\AK{Are there any current analyses with a central jet veto in VBS? The experimentalists might take this statement a little too literal.} On the one hand, in order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable; on the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
 VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data. \AK{Include some references here?}
